Periodic Traveling Waves in an Integro-Difference Equation With a Nonmonotone Growth Function and Strong Allee Effect

### 2021-07-04 paper_draftver7.tex
updates
blossoming soln

By Lemma \ref{lemmamain}, the wave functions $w_1(x)$ and $w_2(x)$ satisfy $Q^2[w_1](x)=w_1(x-2c^*)$ and $Q^2[w_2](x)=w_2(x-2c^*)$. As a result, they can be used to construct a solution to the integral recurrence which spreads in space with a mean speed of $c^*$ and alternates in shape every two time steps. This is formalized in the following theorem:

@@

@@

\subsection{Compactly supported kernels}

The next theorem concerns the spreading behavior of solutions to the IDE \eqref{q} with compactly supported initial data. We also assume the disperal kernel is compactly supported.% First, we introduce a useful lemma for compactly supported kernels.

%\begin{lem}\label{lemma2}
%Suppose $k(x)$ is compactly supported on $[-\sigma,\sigma]$ for some $\sigma>0$, and let $u \in C(\mathbb R)$. If $u(x)=0$ for $x \in \Omega \subseteq \mathbb R$, then $(k*u)(x)=0$ if $d(x,\mathbb R\setminus\Omega):= \inf \{d(x,y) \mid y\in \mathbb R\setminus\Omega\}$.
%\end{lem}
%
%\begin{proof}
%Let $x\in\Omega$ such that $d(x,\mathbb R\setminus\Omega) \geq \sigma$. Then $(x-\sigma,x+\sigma)$ is a subset of  $\Omega$ on which $u$ vanishes; thus $ (k*u)(x) = \int_{x-\sigma}^{x+\sigma}k(x-y)u(y)\,dy = 0.$
%\end{proof}

\begin{lem}
Suppose $k(x)$ has support $[-\sigma,\sigma]$ and $c^*\geq0$.

\begin{enumerate}
\item If $r \geq \sigma - x_\b$, then $Q[w_1](|x|-r) = w_2(|x|-r-c^*)$.

\item 
\end{enumerate}
\end{lem}

\begin{proof}
\begin{enumerate}
\item 
Let $u(x) = w_1(|x|-r)$. Assuming $r+x_\b \geq 0$, i.e. $r \geq -x_\b$, we have $g(u(x)) = \m$ for $|x| < r+x_\b$, $g(u(x)) = 1$ for $r+x_\b \leq x \leq r+x_\a$, and $g(u(x)) = 0 $ for $x > r+x_\a$. Note that $g(u(x)) = g(w_1(x-r))$ for $x \geq -r-x_\b$. Applying the convolution operator, $Q[u](x) = Q[w_1](x-r)$ for $x \geq -r-x_\b+\sigma$. Assuming $r \geq \sigma - x_\b$, this shows $Q[u](x) = w_2(x-r-c^*)$ for $x \geq 0$. By the symmetry of $k(x)$, we have $Q[u](x) = w_2(|x|-r-c^*)$.

\item 
Let $u(x) = w_2(|x|-r)$. Assuming $r+x_\b \geq 0$, i.e. $r \geq -x_\b$, we have $g(u(x)) = \m$ for $|x| < r+x_\b$, $g(u(x)) = 1$ for $r+x_\b \leq x \leq r+x_\a$, and $g(u(x)) = 0 $ for $x > r+x_\a$. Note that $g(u(x)) = g(w_1(x-r))$ for $x \geq -r-x_\b$. Applying the convolution operator, $Q[u](x) = Q[w_1](x-r)$ for $x \geq -r-x_\b+\sigma$. Assuming $r \geq \sigma - x_\b$, this shows $Q[u](x) = w_2(x-r-c^*)$ for $x \geq 0$. By the symmetry of $k(x)$, we have $Q[u](x) = w_2(|x|-r-c^*)$.
\end{enumerate}
\end{proof}

\begin{thm} Suppose $k(x)$ has compact support and $\sup_{x\in\mathbb R}w_2(x)\leq \b$. Let $u_0(x)$ be non-negative and continuous function, and suppose $\a \leq u_0(x) \leq \b$ for $|x| \leq r$ and $u_0(x) < \a$ otherwise, for some $r > 0$. If $r$ is sufficiently large and $c^* \geq 0$, then the sequence $(u_n)_{n=0}^{\infty}$ defined by $u_{n+1}=Q[u_n]$, $n\geq 0$ spreads with speed $c^*$, i.e.
\begin{equation}
\lim_{n\to\infty}\inf_{|x|<n(c^*-\epsilon)}u_{n}(x)>0,
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{|x|>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
for all $\epsilon>0$.
\end{thm}

\begin{thm} Suppose $k(x)$ has compact support and $||w_2||_\infty \leq \b$. If $c^* \geq 0$ and $r$ is sufficiently large, then equation \eqref{q} has a periodic spreading solution given by
\begin{equation} \label{compactptw}
u_n(x) = \begin{cases}
w_1(|x|-r-nc^*) & n \text{ even} \\
w_2(|x|-r-nc^*) & n \text{ odd}
\end{cases}
\end{equation}
for all $n\geq 0$.
\end{thm}

\begin{proof}
As in the proof of Theorem \ref{theorem1}, it suffices to show $u_{n+1}=Q[u_n]$ for all $n\geq 0$. If $n$ is even, we have $u_n(x)=w_1(|x|-r-nc^*)$. Assuming $r \geq \sigma$ and $c^* \geq 0$, this implies $g(u_n(x)) = 0$ for $|x| > r+nc^*+x_\a$, $g(u_n(x))=1$ for $r+nc^*+x_\b\leq |x|\leq r+nc^*+x_\a$, and $g(u_n(x))=\m$ for $|x|<r+nc^*+x_\b$. Note that $g(u_n(x)) = g(w_1(x-r-nc^*))$ for $x \geq -r-nc^*-x_\b$. Applying the convolution operator, we have $Q[u_n](x) = Q[w_1](x-r-nc^*)$ for $x \geq -r-nc^*-x_\b+\sigma$. Assuming $r \geq \sigma - x_\b$, it follows that $Q[u_n](x) = Q[w_1](x-r-nc^*)$ for $x \geq 0$. By hypothesis (H2), $k(x)$ is an even function, thus $Q[u_n](-x) = Q[u_n](x)$, hence $Q[u_n](x) = Q[w_1](|x|-r-nc^*)$ for all $x \in \mathbb R$, i.e. $u_{n+1}(x) = w_2(|x|-r-nc^*-c^*)$. Note that the preceding calculation requires $r \geq \max\{\sigma,\sigma-x_\b\}$; a sufficient condition is $r \geq 2\sigma$.

If $n$ is odd, then $u_n(x)=w_2(|x|-r-nc^*)$, and for $r\geq0$ and $c^*\geq 0$ it follows that $g(u_n(x)) = 0$ for $|x| > r+nc^*+c^*$ and $g(u_n(x))=1$ for $|x| \leq r+nc^*+c^*$.

UNFIN
\end{proof}

%
%We apply induction on $n$. For the basis case, it is sufficient to show $u_1(x) = w_1(|x|-r)$ for all $x$. It follows from the definition of $u_0(x)$ that $g(u_0(x))=1$ for $|x|\leq r$ and $g(u_0(x))=0$ otherwise. Assuming $r \geq \frac{\sigma}{2}$, we have $u_1(x)=Q[u_0](x)=w_1(x-r)$ for $x\geq 0$. By hypothesis (H2), $k(x)$ is an even function, so $u_1(x)=u_1(-x)$, i.e. $u_1(x)=w_1(-x+r)=w_1(x-r)$ for $x<0$. This proves $u_1(x)=w_1(|x|-r)$ for all $x$.
%
%For the inductive step, suppose equation \eqref{compactptw} holds for some $n\geq0$. If $n$ is even, the inductive hypothesis yields $u_{n+1}(x)=w_1(|x|-r-nc^*)$.
%\end{proof}
%
%\begin{proof}
%By the translation invariance property of $Q$, we may assume without loss of generality $A=[-r,r]$, for some $r>0$. It follows that $u_0(x)<\a$ for $|x|>r$ and $\a\leq u_0(x)\leq\b$ for $|x| \leq r$. Furthermore, there exists $\sigma>0$ such that $k(x)>0$ if and only if $|x|<\sigma$. Assuming $r \geq \frac{\sigma}{2}$, we have
%\begin{equation} \begin{aligned}
%u_1(x) = Q[u_0](x) &= \int_{-r}^{r} k(x-y) \, dy \\
%&= \begin{cases}
%w_1(x-r) & x \geq 0 \\
%w_1(-x+r) & x< 0
%\end{cases}
%\end{aligned}  \end{equation}
%
%Let $x_\a$ and $x_\b$ be as defined in Lemma \ref{lemma1}. Then, applying the growth function, we have
%\begin{equation} g(u_1(x)) = \begin{cases}
%0 & x<-r-x_\a \\
%1 & -r-x_\a \leq x \leq -r-x_\b \\
%m & -r-x_\b < x < r+x_\b \\
%1 & r+x_\b \leq x \leq r+x_\a \\
%0 & x > r+x_\a
%\end{cases}
%\end{equation}
%Note that $g(u_1(x))=g(w_1(x))$ for all $x >-r-x_\b$. Thus $Q[u_1](x)=(k*(g\circ u_1))(x)=(k*(g\circ w_1))(x)=Q[w_1](x)$ for all $x>-r-x_\b+\sigma$. If we assume $r \geq x_\b+\sigma$, then $u_2(x)=Q[u_1](x)=Q[w_1](x-r)=w_2(x-r)$ for $x>0$.  By symmetry of the integro-difference operator, $Q[u_1](x) = Q[w_1](-x+r)$ for $x<0$. Thus,
%
%\begin{equation}
%u_2(x) = Q[u_1](x) = \begin{cases}
%w_2(x-r) & x \geq 0 \\
%w_2(-x+r) & x < 0 
%\end{cases}
%\end{equation}
%Since $c^*\geq 0$, we have
%\begin{equation}
%g(u_2(x)) = \begin{cases}
%0 & x < -r-2c^* \\
%1 & -r-2c^* \leq x \leq r+2c^* \\
%0 & x > r+2c^* 
%\end{cases}
%\end{equation}
%
%Thus,
%\begin{equation}
%u_3(x) = Q[u_2](x) = \begin{cases}
%w_1(x-r-2c^*) & x \geq 0 \\
%w_1(-x+r+2c^*) & x < 0 
%\end{cases}
%\end{equation}
%
%The preceding argument can be repeated inductively to obtain
%\begin{equation}
%u_{2n+1}(x) = w_1(|x|-r-2nc^*)
%\end{equation}
%and
%\begin{equation}
%u_{2n+2}(x) = w_2(|x|-r-2nc^*)
%\end{equation}
%for all $n\geq 0$.
%\end{proof}

@@

\subsection{Variability in wave speed}

@@

\begin{ex}
The Laplace kernel,
\begin{equation}\label{laplacekernel}
k(x) = \frac{1}{2} e^{-|x|}
\end{equation}

@@

\begin{equation} \label{w1laplace}
w_1(x) =   \begin{cases} 
1 - \frac{1}{2}e^{x} & x \leq 0 \\
\frac{1}{2}e^{-x} & x > 0
\end{cases}
\end{equation}
and
%\begin{equation}\label{laplacew2}
%w_2(x) = \begin{cases}
%m - \frac{1}{2} e^{x - \alpha} + \frac{1-\m}{2} e^{x - \beta} & x < \beta \\
%1 - \frac{1}{2} e^{x-\alpha} - \frac{1-\m}{2} e^{-(x-\beta)} & \beta \leq x \leq \alpha \\
%\frac{1}{2} e^{-(x-\alpha)} - \frac{1-\m}{2} e^{-(x-\beta)} & x > \alpha
%\end{cases}
%\end{equation}
%\begin{equation} \label{w2laplace}
%w_2(x) = \begin{cases}
%m + C_1 e^x & x < x_\b \\
%1 - C_2 e^x - C_3 e^{-x} & x_\b \leq x \leq x_\a \\
%C_4 e^{-x} & x_\a < x 
%\end{cases} \end{equation}
%or equivalently

\begin{equation} \label{w2laplace}
w_2(x) = \begin{cases}
m + \left( \frac{1-\m}{2}e^{-x_\b} - \frac{1}{2}e^{-x_\a} \right) e^x & x < x_\b \\
1 - \left( \frac{1}{2} e^{-x_\a} \right) e^x - \left( \frac{1-\m}{2} e^{x_\b} \right)  e^{-x} & x_\b \leq x \leq x_\a \\
\left( \frac{1}{2} e^{x_\a} - \frac{1-\m}{2} e^{x_\b} \right) e^{-x} & x_\a < x 
\end{cases} \end{equation}

where

@@

x_\a = \begin{cases} -\ln(2\a) & \a\leq \frac{1}{2} \\ \ln(2-2\a) & \a > \frac{1}{2} \end{cases}, \quad
x_\b = \begin{cases} -\ln(2\b) & \b\leq \frac{1}{2} \\ \ln(2-2\b) & \b > \frac{1}{2} \end{cases} 

@@

\frac{1}{2}x_\a  + \frac{1}{2}\log \left( \frac{1 - (1-\m) e^{x_\b-x_\a}}{2\a} \right) & \a \leq w_2(x_\a) \\
\frac{1}{2}x_\a + \frac{1}{2} \log \left( 1-\a + \sqrt{(1-\a)^2 - (1-\mu) e^{x_\b-x_\a}} \right) & w_2(x_\a) < \a < w_2(x_\b) \\
\frac{1}{2}x_\a + \frac{1}{2} \log \left( \frac{2(\a-\m)}{(1-\m)e^{x_\a-x_\b}-1} \right) & \a \geq w_2(x_\b)

>>>>

By Lemma \ref{lemmamain}, the wave functions $w_1(x)$ and $w_2(x)$ satisfy $Q^2[w_1](x)=w_1(x-2c^*)$ and $Q^2[w_2](x)=w_2(x-2c^*)$. As a result, they can be used to construct a solution which spreads in space with a mean speed of $c^*$ and alternates in shape every two time steps.

@@

We now introduce an additional hypothesis for the spread of solutions with compactly supported initial condition.

Hypothesis 2. For sufficiently large $r_b$, with $r_a > r_b > 0$ and $\lambda \in (0,1)$, the equation

$$ k(x+r_a) - k(x-r_a) - \lambda k(x+r_b) + \lambda k(x-r_b) $$

has at most 3 solutions.

@@

\subsection{Spreading solutions}

In this section, we consider how the periodic traveling wave solutions in the preceding results correspond to solutions with a compactly supported initial population density. For simplicity, we restrict our attention to initial conditions where the maximum population value does not exceed the overcompensation threshold.

Let $\mathbf 1_{[-r,r]}$ denote the indicator function on $[-r,r]$. Suppose $u_0$ is a continuous function with compact support, and suppose $u_0$ satisfies $\a \leq u_0(x) \leq \b$ for $|x| \leq r$ and $u_0(x) < \a$ otherwise. It follows that
$$ g(u_0(x)) = \mathbf 1_{[-r_0,r_0]}(x) $$
We have
$$ Q[u_0](x) = (k*(g \circ u_0))(x) = K(x+r_0) - K(x-r_0) $$
Assuming $r_0$ is large enough so that $ Q[u_0](x) = 2K(r_0)-1 \geq \b$, it follows that
$$ g(Q[u_0](x)) = \begin{cases}
n_1 & |x| < r_1 \\
n_2 & r_1 \leq |x| \leq r_2 \\
0 & |x| > r_2
\end{cases} $$
It follows that
$$ Q^2 [u_0] (x) = n_2 \int_{-r_2}^{r_2} k(x-y)\,dy - (n_2-n_1) \int_{-r_1}^{r_1} k(x-y) \, dy $$
$$ = n_2 K(x+r_2) - n_2K(x-r_2) - (n_2-n_1) K(x+r_1) + (n_2-n_1) K(x-r_1) $$
Taking the derivative, we find
$$ \frac{d}{dx} Q^2 [u_0](x) = n_2 k(x+r_2) - n_2k(x-r_2) - (n_2-n_1) k(x+r_1) + (n_2-n_1) k(x-r_1) $$

We now require the following hypothesis on the disperal kernel:

(H5) For all $r_2>r_1>0$ and $\lambda \in (0,1)$, the equation
$$ k(x+r_2) - k(x-r_2) - \lambda k(x+r_1) + \lambda k(x-r_1) = 0 $$
has at most 3 solutions.

If hypothesis (H5) holds, we can then deduce that $\frac{d}{dx}Q^2[u_0]$ has at most 3 zero-crossings, hence, $Q^2[u_0]$ has at most 3 turning points. This leaves only two cases: either $Q^2[u_0]$ has a single turning point at $x=0$, or $Q^2[u_0]$ has 3 turning points, one at $x=0$, and two diametrically opposed turning points $+t$ and $-t$. In both cases, there is a unique $r_2>0$ satisfying $Q^2[u_0](x) \geq \a$ for $|x| \leq r_2$ and $Q^2[u_0](x) < \a$ otherwise. If $r_2 < r$, then the conclusion does not follow. But if $r_2 \geq r$, then the

The next theorem concerns the spreading behavior of solutions to the IDE \eqref{q} with compactly supported initial data.

@@

\subsection{Blossoming solution}

In addition to the behavior mentioned earlier, we noticed the phenomena of blossoming solution. Whenever

@@

\begin{ex}
The Laplace kernel is given by
\begin{equation}\label{laplacekernel}
k(x) = \frac{1}{2} e^{-|x|}
\end{equation}

@@

\begin{equation} \label{w1laplace}
w_1(x) =   \begin{cases} 
n_2 - \frac{1}{2} n_2 e^{x} & x \leq 0 \\
\frac{1}{2} n_2 e^{-x} & x > 0
\end{cases}
\end{equation}
and
%\begin{equation}\label{laplacew2}
%w_2(x) = \begin{cases}
%m - \frac{1}{2} e^{x - \alpha} + \frac{1-\m}{2} e^{x - \beta} & x < \beta \\
%1 - \frac{1}{2} e^{x-\alpha} - \frac{1-\m}{2} e^{-(x-\beta)} & \beta \leq x \leq \alpha \\
%\frac{1}{2} e^{-(x-\alpha)} - \frac{1-\m}{2} e^{-(x-\beta)} & x > \alpha
%\end{cases}
%\end{equation}
%\begin{equation} \label{w2laplace}
%w_2(x) = \begin{cases}
%m + C_1 e^x & x < x_\b \\
%1 - C_2 e^x - C_3 e^{-x} & x_\b \leq x \leq x_\a \\
%C_4 e^{-x} & x_\a < x 
%\end{cases} \end{equation}
%or equivalently

\begin{equation} \label{w2laplace}
w_2(x) = \begin{cases}
\m + \left( \frac{1}{2}(\mtwo-\m)e^{-x_\b} - \frac{1}{2}\mtwo e^{-x_\a} \right) e^x & x < x_\b \\
\mtwo - \frac{1}{2} \mtwo e^{-x_\a} e^x - \frac{1}{2} (\mtwo-\m) e^{x_\b}   e^{-x} & x_\b \leq x \leq x_\a \\
\left( \frac{1}{2} \mtwo e^{x_\a} - \frac{1}{2}\mtwo(\mtwo-\m) e^{x_\b} \right) e^{-x} & x_\a < x 
\end{cases} \end{equation}

with $x_a = F(a)$, $x_b=F(b)$, and

@@

F(p) = \begin{cases} -\ln(2p) &p \leq \frac{1}{2} \\ \ln(2-2p) & p > \frac{1}{2} \end{cases}

@@

\frac{1}{2} \log \left( \frac{1}{2\a} \mtwo e^{x_\a} - \frac{1}{2\a}(\mtwo-\m)e^{x_\b} \right) & \a \leq w_2(x_\a) \\
\frac{1}{2} \log \left( \frac{1}{\mtwo} e^{-x_\a} \left( \m - \a + \sqrt{(\m-\a)^2-\mtwo(\mtwo-\m)e^{x_\b-x_\a}}\right)\right) & w_2(x_\a) < \a < w_2(x_\b) \\
\frac{1}{2} \log \left( \frac{2(\a-\m)}{(\mtwo-\m)\exp(-x_\b)-\mtwo \exp(-x_\a)}\right) & \a \geq w_2(x_\b)
